---
title: "A2S exploratory analysis"
author: "Renato Russo"
output:
  html_document:
    df_print: paged
    code_folding: hide
---

This is a document with ongoing exploratory analysis of data from the A2S project.

```{r Loading data, echo=FALSE}
load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_events.Rda")
a2s_events <- a2s_events

load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_progress.Rda")
a2s_progress <- a2s_progress

load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_logs.Rda")
a2s_logs <- a2s_logs

load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_students.Rda")
a2s_students <- a2s_students

library(summarytools)
library(tidyverse)
```

## Exploratory analysis
### Table: Absolute frequency of journeys
#### journeyId = 9: diffusion
#### journeyId = 44: wildfires
```{r}
table(a2s_events$journeyId)
```

### Table: frequency of tasks
#### taskId = 6 and 451: part of the diffusion journey
#### taskId = 41 and 408: part of the wildfires journey
```{r}
table(a2s_events$taskId)

```
### Table: Absolute frequency of block types
```{r}
table(a2s_events$blockType)
```

### Table: Block type per journey (percentages add in the horizontal)
```{r}
ctable(
  x = a2s_events$blockType,
  y = a2s_events$journeyId
)
```

```{r}
### Block types by journey ----------------------------------------------------
# a table of block types by journey
blocktype_journey <- (table(a2s_events$blockType, a2s_events$journeyId))
colnames(blocktype_journey)[colnames(blocktype_journey) == "9"] <- "diffusion"
colnames(blocktype_journey)[colnames(blocktype_journey) == "44"] <- "wildfires"
as.data.frame(blocktype_journey)

table(blocktype_journey)
```
```{r}
# a table of the block types for journey "diffusion"
diffusion_blocktypes <- as.data.frame(blocktype_journey[, 'diffusion'])
View(diffusion_blocktypes)
```

# a table of the block types for journey "wildfires"
wildfires_blocktypes <- as.data.frame(blocktype_journey[, 'wildfires'])
View(wildfires_blocktypes)

### Action types by journey
# a table of action types by journey
action_journey <- (table(a2s_events$actionType, a2s_events$journeyId))
colnames(action_journey)[colnames(action_journey) == "9"] <- "diffusion"
colnames(action_journey)[colnames(action_journey) == "44"] <- "wildfires"
as.data.frame(action_journey)
action_journey

# a table of the action types for journey "diffusion"
diffusion_actiontypes <- as.data.frame(action_journey[, 'diffusion'])
View(diffusion_actiontypes)

# a table of the block types for journey "wildfires"
wildfires_actiontypes <- as.data.frame(action_journey[, 'wildfires'])
View(wildfires_actiontypes)
```