---
title: "A2S exploratory analysis"
author: "Renato Russo"
output:
  html_document:
    df_print: paged
    code_folding: hide
---

This is a document with ongoing exploratory analysis of data from the A2S project.

```{r Loading data, echo=FALSE}
load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_events.Rda")
a2s_events <- a2s_events

load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_progress.Rda")
a2s_progress <- a2s_progress

load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_logs.Rda")
a2s_logs <- a2s_logs

load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_students.Rda")
a2s_students <- a2s_students

load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_logs_clean.Rda")
a2s_logs_clean <- a2s_logs_clean

library(summarytools)
library(tidyverse)
library(ggplot2)
```

## Exploratory analysis - "events"
### Table: Absolute frequency of journeys
#### journeyId = 9: diffusion
#### journeyId = 44: wildfires
```{r}
table(a2s_events$journeyId)
```

### Table: frequency of tasks
#### taskId = 6 and 451: part of the diffusion journey
#### taskId = 41 and 408: part of the wildfires journey
```{r}
table(a2s_events$taskId)

```
### Table: Absolute frequency of block types
```{r}
table(a2s_events$blockType)
```

### Table: Block type per journey (percentages add in the horizontal)
```{r}
ctable(
  x = a2s_events$blockType,
  y = a2s_events$journeyId
)
```

### Table: Block type per journey
```{r}
### Block types by journey ----------------------------------------------------
# a table of block types by journey
blocktype_journey <- (table(a2s_events$blockType, a2s_events$journeyId))
colnames(blocktype_journey)[colnames(blocktype_journey) == "9"] <- "diffusion"
colnames(blocktype_journey)[colnames(blocktype_journey) == "44"] <- "wildfires"
as.data.frame(blocktype_journey)
blocktype_journey

```
### Table: Block type frequency (journey: diffusion)
```{r}
# a table of the block types for journey "diffusion"
diffusion_blocktypes <- as.data.frame(blocktype_journey[, 'diffusion'])
diffusion_blocktypes

```
### Table: Block type frequency (journey: "wildfires")
```{r}
# a table of the block types for journey "wildfires"
wildfires_blocktypes <- as.data.frame(blocktype_journey[, 'wildfires'])
wildfires_blocktypes
```

## Exploratory analysis - "logs" (video-viewing data)
### Table: video frequency
```{r}
table(a2s_logs_clean$Video)
```
### Table: Video frequency
```{r}
freq(a2s_logs_clean$Video)
```

### Table: Frequency of actions by video
```{r}
ctable(
  x = a2s_logs_clean$Video,
  y = a2s_logs_clean$Action
)
```

### Table: Frequency of actions by user
```{r}
ctable(
  x = a2s_logs_clean$Action,
  y = a2s_logs_clean$userId
)
```

```{r}
ggplot(a2s_logs_clean, aes(x = Time)) + geom_density(stat = "bin", binwidth = 3600, 
    colour = "black", fill = "darkgreen") + ylab("Requests/hour") + xlab("Time")
```