---
title: "Individual Coding Exercise - Missing Data Imputation"
output: html_document
df_print: paged
author: Renato Russo
date: April, 15, 2022
---

# Getting and preparing the data
```{r}
library(datasets)
data(iris)
head(iris)
```

# MCAR
```{r}
iris_MCAR <- iris

missing_fraction <- 0.3 #fraction of missing variables 
N <- nrow(iris_MCAR)

is.na(iris_MCAR[sample.int(N, missing_fraction*N), "Sepal.Length"]) <- TRUE

iris_MCAR
```

# MAR
##logit and inverse logistic functions
```{r}
logit <- function(x){log(x/(1 - x))}
invlogit <- function(x){1/(1 + exp(-x))}
```

##generate a Bernoulli distribution where p is 0.3
```{r}
m.SW <- mean(iris$Sepal.Width)
pmiss <- invlogit(iris$Sepal.Width -m.SW + logit(0.3))
hist(pmiss) 

iris_MAR <- iris
N <- nrow(iris_MAR)
is.na(iris_MAR$Sepal.Length[runif(N) < pmiss]) <- TRUE
```

#Deletion methods
##Listwise deletion
iris_MCAR_listwise <- iris_MCAR[complete.cases(iris_MCAR),]
nrow(iris_MCAR_listwise)

##Pairwise deletion
mean(iris_MCAR$Sepal.Length, na.rm = TRUE)
sd(iris_MCAR$Sepal.Length, na.rm = TRUE)

#Single imputation
##Simple imputation
```{r}
iris_MCAR_mean <- iris_MCAR
mi <- mean(iris_MCAR$Sepal.Length, na.rm = TRUE)
iris_MCAR_mean$Sepal.Length[is.na(iris_MCAR$Sepal.Length)] <- mi
iris_MCAR_mean
```

##Regression imputation/conditional mean imputation
```{r}
iris_MAR_regression <- iris_MAR
lm_model <- lm(Sepal.Length ~ Sepal.Width, iris_MAR_regression)
summary(lm_model)
```

###Broadcasting the Beta and the intercept to all missing Sepal length
```{r}
iris_MAR_regression$Sepal.Length[is.na(iris_MAR$Sepal.Length)] <- -0.08 * iris_MAR_regression$Sepal.Width[is.na(iris_MAR$Sepal.Length)] + 6.085

iris_MAR_regression
```
##Hotdeck imputation or nearest neighbors
```{r}
#install.packages("missMethods")
library(missMethods)
iris_MAR_hotdeck <- iris_MAR
impute_hot_deck_in_classes(iris_MAR_hotdeck, c("Sepal.Width", "Petal.Length", "Petal.Width", "Species"))

```