---
title: "Individual Coding Exercise - Knowledge Inference"
output: html_document
df_print: paged
author: Renato Russo
date: April, 16, 2022
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Getting and preparing the data:
```{r}
ADL <- read.table(
  "http://publicifsv.sund.ku.dk/~kach/scaleval_IRT/ADL.txt", 
  sep=' ', 
  header = TRUE,
  na.strings = '.')
names(ADL)

comp <- complete.cases(ADL)
ADL.comp <- ADL[comp,]
items <- ADL.comp[,-1]
```

Installing the ltm package (used for "Analysis of multivariate dichotomous and polytomous data using latent trait models under the Item Response Theory approach")
```{r}
install.packages("ltm", repos="http://cran.rstudio.com/")
library(ltm)
```

Overview of items and item-total correlations. Fitting 1PL (or Rasch model) using the code below:

```{r}
fit1 <- rasch(items)
coef(fit1, prob = T, order = T)
```

Getting all ICCs in a plot:
```{r}
plot(fit1, type = "ICC")
```
Evaluating model fit in dichotomus IRT models using ltm()
```{r}
GoF.rasch(fit1, B=99, seed=1862)
```

Evaluating item fit in dichotomus IRT models using ltm()
```{r}
if1 <- item.fit(fit1, 
                  G = 2, 
                  FUN = mean, 
                  simulate.p.value = TRUE, 
                  B = 99)
if1
```

Item Information Functions
```{r}
plot(fit1, type = "IIC")
```

Test Information functions
```{r}
plot(fit1, type = "IIC", items=0)
```
