---
title: "A2S exploratory analysis"
author: "Renato Russo"
output:
  html_document:
    df_print: paged
    code_folding: hide
---

This is a document with ongoing exploratory analysis of data from the A2S project. The project is part of a larger research effort to create a constructionist environment for learning of "real world data for sustained investigations."
The data consists of a log file and a table of video viewing data.

```{r Loading data, echo=FALSE}
#load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_events.Rda")
load("C:\\Users\\Renato\\Desktop\\HUDK-4051\\A2S\\a2s_events.Rda")
a2s_events <- a2s_events

#load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_progress.Rda")
load("C:\\Users\\Renato\\Desktop\\HUDK-4051\\A2S\\a2s_progress.Rda")
a2s_progress <- a2s_progress

#load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_logs.Rda")
load("C:\\Users\\Renato\\Desktop\\HUDK-4051\\A2S\\a2s_logs.Rda")
a2s_logs <- a2s_logs

#load("/Users/renatorusso/Desktop/Ed.D./HUDK-4051/A2S/a2s_students.Rda")
load("C:\\Users\\Renato\\Desktop\\HUDK-4051\\A2S\\a2s_students.Rda")
a2s_students <- a2s_students

library(summarytools)
library(tidyverse)
```

## Exploratory analysis of the logfile
### Table: Absolute frequency of journeys
#### journeyId = 9: diffusion
#### journeyId = 44: wildfires
```{r}
table(a2s_events$journeyId)
```

### Table: frequency of tasks
#### taskId = 6 and 451: part of the diffusion journey
#### taskId = 41 and 408: part of the wildfires journey
```{r}
table(a2s_events$taskId)

```
### Table: Absolute frequency of block types
```{r}
table(a2s_events$blockType)
```
### Table: Absolute frequency of block types after categorization
```{r}
table(a2s_events$blockTypeCategory)
```

### Table: Block type per journey (percentages add in the horizontal)
```{r}
ctable(
  x = a2s_events$blockTypeCategory,
  y = a2s_events$journeyId
)
```

### Block types by journey
# a table of block types by journey
```{r}
blocktype_journey <- (table(a2s_events$blockTypeCategory, a2s_events$journeyId))
colnames(blocktype_journey)[colnames(blocktype_journey) == "9"] <- "diffusion"
colnames(blocktype_journey)[colnames(blocktype_journey) == "44"] <- "wildfires"
as.data.frame(blocktype_journey)

table(blocktype_journey)
```

### a table of the block types for journey "diffusion"
```{r}
diffusion_blocktypes <- as.data.frame(blocktype_journey[, 'diffusion'])
diffusion_blocktypes
```

### a table of the block types for journey "wildfires"
```{r}
wildfires_blocktypes <- as.data.frame(blocktype_journey[, 'wildfires'])
wildfires_blocktypes
```
### Action types by journey
# a table of action types by journey
```{r}
action_journey <- (table(a2s_events$actionType, a2s_events$journeyId))
colnames(action_journey)[colnames(action_journey) == "9"] <- "diffusion"
colnames(action_journey)[colnames(action_journey) == "44"] <- "wildfires"
as.data.frame(action_journey)
action_journey
```

### a table of the action types for journey "diffusion"
```{r}
diffusion_actiontypes <- as.data.frame(action_journey[, 'diffusion'])
diffusion_actiontypes
```

### a table of the block types for journey "wildfires"
```{r}
wildfires_actiontypes <- as.data.frame(action_journey[, 'wildfires'])
wildfires_actiontypes
```

## Exploratory analysis of the video data
### Frequency table for each video
```{r}
library(summarytools)
freq(a2s_logs$Video)
```

### Frequency of actions by video
```{r}
ctable(
  x = a2s_logs$Video,
  y = a2s_logs$Action
)
```

```{r}
ggplot(a2s_logs) +
  geom_bar(mapping = aes(x = a2s_logs$Video), binwidth = 0.5)

```
```{r}
ggplot(a2s_logs) +
  geom_bar(mapping = aes(a2s_logs$Video, fill = a2s_logs$Action))

```


### Frequency table of action by user
```{r}
ctable(
  x = a2s_logs$Action,
  y = a2s_logs$userId
)
```

### Frequency table of video by user
```{r}
ctable(
  x = a2s_logs$Video,
  y = a2s_logs$userId
)
```